---
title: "Card Creation"
output: html_document
date: "2023-03-16"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## How convenience is created

This document illustrates convenience functionality to help with card creation.
The main idea how to make card creation more convenient is as follows. The Metabase
client works a lot with ids for robustness reasons. As a result direct usage of the client
involves a lot of id look ups in order to create a query. The convenience functionality
aims to spare the user these manual lookups but instead looks all this up in the
backgroud. In order to do that a Metabase client typically needs to be supplied as an
argument.

On the one hand this creates more convenience, but on the other hand it is
potentially more error prone, as there might be non-unique names (even though
ids would of course be unique). If such cases are encountered it might not be possible
to use the convenience functionality.

For the remaining examples in this document we created a MetabaseClient `mc`. And
a collection `card_creation_example` (if it doesn't already exist). All the example
cards are created within this collection. As the simplest example to illustrate
that card creation was successful we will run the cards and show a sample
of the output as this gives a tangible result.

# Examples

## Simple query creation

```{r load_package and create client, echo=FALSE,warning=FALSE,message=FALSE}
devtools::load_all()
require("dplyr")

mc <- MetabaseClient()
```

```{r card_creation_collection, echo=FALSE}
collection_name <- "card_creation_example"

collections <- mc$get_collections()
if (!(collection_name %in% (collections %>% pull(name)))) {
  mc$create_collection(collection_name)
}
```

```{r create_card}
card_id <- create_select_all_card(mc, "My new card", "card_creation_example", "ORDERS")
mc$run_card(card_id) %>% head()
```
