---
title: "Reference Class Client"
output: html_document
date: "2022-11-27"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Reference Class Client for Metabase 

Reference classes (RC) allow the definition of objects that have an internal state
that can be modified by methods. In the context of Metabase this allows us
to write a client that encapsulates the session behavior of the API. This way
users of the client do not have to deal with sessions directly.

```{r import client}
devtools::load_all()
```

A client can be instantiated from an RC. These classes also allow for more
complex initialization which enables us to encapsulate loading environment
variables. In order to create a valid client instance the following code
requires the environment variables `METABASE_USER`, `METABASE_PWD` and
`METABASE_URL`.

```{r instanciate}
mc <- MetabaseClient()
```

Given an instance, interaction with the API can begin immediately with all
session necessities happening in the background.

```{r get_databases}
mc$get_databases()
```
