---
title: "Reference Class Client"
output: html_document
date: "2022-11-27"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Reference Class Client for Metabase 

Reference classes (RC) allow the definition of objects that have an internal state
that can be modified by methods. In the context of Metabase this allows us
to write a client that encapsulates the session behavior of the API. This way
users of the client do not have to deal with sessions directly.

```{r import client}
devtools::load_all()
```

A client can be instantiated from an RC. These classes also allow for more
complex initialization which enables us to encapsulate loading environment
variables. In order to create a valid client instance the following code
requires the environment variables `METABASE_USER`, `METABASE_PWD` and
`METABASE_URL`.

```{r instanciate}
mc <- MetabaseClient()
```

Given an instance, interaction with the API can begin immediately with all
session necessities happening in the background.

```{r get_databases}
mc$get_databases()
```

In the rest of the document we will illustrate further examples using the `Sample Database` as this should always be present. It likely has the id 1, but to be on the safe side we extract the id programatically and store it in a variable.

```{r import_dplyr, echo=FALSE,warning=FALSE,message=FALSE}
require("dplyr")
```

```{r sample_database-id}
sample_db_id <- mc$get_databases() %>%
  filter(name == "Sample Database") %>%
  pull(id)
```

With the help of a database id we can habe a look at all the collections that are available for
this database.

```{r all_collections}
collections <- mc$get_collections()
collections
```

By default there is only our personal collection and the root collection.
For the further illustrations we'll create a specific collection if it does not exist yet.

```{r metabase_example_collection}
example_collection <- "Metabase Example Collection"
personal_collection_id <- as.integer((collections %>% filter(grepl("Personal Collection", name)) %>% pull(id))[1])

if (!(example_collection %in% (collections %>% pull(name)))) {
  collection_details <- mc$create_collection(example_collection, parent_collection_id = personal_collection_id)
  example_collection_id <- collection_details$id
} else {
  example_collection_id <- collections %>%
    filter(name == example_collection) %>%
    pull(id)
}
mc$get_collections()
```

As we can see here we created a new collection. We also stored its ID for future reference.
At this stage we can also check the web UI to see that collection was actually created
there as well.

Currently our new collection is empty and therefore not very interesting. Nonetheless
we can actually check that there are no items in the collection

```{r collection_items}
mc$get_items(collection_id = example_collection_id)
```

While we could add child collections to our parent collection if we so desired, this
is of course not the purpose of Metabase. Instead we we want to actually add cards that
operate on the sample data. This is done with the `create_card` method of the client,
which takes three arguments. The name of the card we want to create, the collection_id
where we want to create it, and the query associated with the card. (Remember you can look
at the client method documentation with `?MetabaseClient`). The first two arguments are straightforward,
but the query argument is a little more difficult. That is why this projects provides helper functions
to create queries. We'll conclude this document by giving a small example, and go into more depth
about query creation in other documents.

In order to create a specific data example, we need to know which data is abvailbe, which we can check with the `get_tables` method.

```{r get_tables}
tables <- mc$get_tables()
order_table_id <- tables %>%
  filter(table_name == "ORDERS") %>%
  pull(table_id)
tables
```

As an example, we pick the `ORDERS` table and already extracted its ID in the example above.
This way we can create a query and a card as follows.
```{r card_creation}
query <- create_select_all_columns(table_id = order_table_id, database_id = sample_db_id)
card_name <- "Our first test card"
card_id <- mc$create_card(card_name, collection_id = example_collection_id, query = query)
```

We can see that the card was created with `get_items`.
```{r card_creation_check}
items <- mc$get_items(example_collection_id)
items
```

We can also run the card to see the data output with `run_card`. 

```{r run_card_example}
mc$run_card(card_id)
```
